---
title: "Mini Project 1"
author: "Prajwal Bhandari"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Front matter
```{r}
# import libraries
# remove(list = ls())
library(tidyverse)
library(readxl)
library(forcats)
# read data
cod <- read_xlsx("CODGames2_mp.xlsx", sheet = 1)
```


## Preliminary 
```{r}
glimpse(cod)
```

In this data, there are 211 observations, each with 27 different attributes. Variables like `map1`, `map2`, `PrimaryWeapom`, and other character variables could be converted into factors for easier analysis. The variable `XPType` could be converted to a numerical value instead of staying as a character variable. We could do some easy math to arrive at an XP multiplier that is applied to each game. Finally, variables like the date could be parsed by functions from the lubridate package that would make working with them easier. Other work like changing all the NA values in `ObjejctiveKills`, `Captures`, `Plants`, and others would make more sense to make the analysis easier and so that we don't have to work with missing data. Also, rename variables to be in snake case. 

We can make the following changes and then begin variable exploration. 
```{r}
cod_working <- cod %>%
  rename(
    map1 = Map1,
    map2 = Map2,
    choice = Choice,
    map_vote = MapVote,
    date = Date,
    full_partial = FullPartial,
    result = Result,
    eliminations = Eliminations,
    deaths = Deaths,
    score = Score,
    damage = Damage,
    total_xp = TotalXP,
    primary_weapon = PrimaryWeapon,
    xp_multiplier = XPType, # call this multiplier, will convert in next command
    did_player_vote = DidPlayerVote,
    game_type = GameType,
    confirms = Confirms,
    denies = Denies,
    objectives = Objectives,
    objective_kills = ObjectiveKills,
    captures = Captures,
    diffuses = Diffuses,
    plants = Plants,
    detonates = Detonates,
    deposits = Deposits,
    time_sec = Time_Sec,
    time_min = Time_Min
  ) %>%
  mutate(  # for now, only mutating variables we think are important
    map1 = as.factor(map1), 
    map2 = as.factor(map2),
    choice = as.factor(choice),
    full_partial = as.factor(full_partial),
    primary_weapon = as.factor(primary_weapon),
    xp_multiplier = if_else(xp_multiplier == '10% Boost', 1.10, 2.10),
    did_player_vote = as.logical(if_else(did_player_vote == 'Yes',1,0)),
  ) %>%
  separate(result, into = c('team_result', 'other_team_result'), sep = '-', extra = 'merge', convert = TRUE) 


# figure out na values for each column
apply(cod_working, MARGIN = 2, FUN = function(x) sum(is.na(x))) # counts total NA in columns
# since NA values are either < 50 or >=200 drop all columns with more than 200 NA values

cod_working <- cod_working %>%
  select(where(~ sum(is.na(.)) < 200))

```




## Part 1 

The first variable we decided to explore is `choice`. It is a categorical variable that represents the map that each game was played on and there is only one missing value for this variable. It's reasonable to state that the count of each variable is the number of games played on that map. 

We see the number of 
```{r}
sum(is.na(cod_working$choice))
```

```{r}
data.frame(cod_working$choice) %>% # convert sngle col into dataframe
  na.omit() %>% # makes sense to omit single observation in one variable
  ggplot(aes(x = cod_working.choice)) +
  geom_bar(fill = 'lightgreen', color = 'black') +
  labs(
    x = 'Map Choice',
    y = 'Games Played ', 
    title = 'Bar Plot of Map Choice and Games Played',
    caption = '* one dropped value'
  ) +
  coord_flip() +
  theme_bw()
```

From the bar chart above, there are a few observations that we can make. The least chosen map is 'Amerika' with 3 picks, the the most chosen are 'Rush', 'Diesel', and 'Crossroads Strike' with 15 picks. Other maps tend to be between 5 and 12 picks. It's interesting to see that the maps are not chosen equally. Further questions could be asked about why some maps are chosen more than others but that is outside the scope of our analysis. 

The third variable that we thought is interesting to explore is the `primary_weapon` variable, which is self explanatory. There are 9 different primary weapons that are used: M16, MP5, AK47, Krig 6, QBZ-83, Pelington 703 FFAR 1, Type 63, and the MG82. Like we saw above, there are no missing values in this variable.

```{r}
ggplot(data = cod_working, mapping = aes(x = fct_reorder(primary_weapon,primary_weapon, .fun = length))) + 
  geom_bar(fill = 'lightgreen', color = 'black') +
  labs(
    x = 'Primary Weapon Used', 
    y = 'Number of Games',
    title = 'Counts for Primary Weapon'
  ) + 
  theme_bw()
```

It's interesting to see the disparity between weapon choice. The M16 was used in 48 games, while weapons like the MG 82, FFAR 1 and AK-47 were used in less than 5 games. Some potential reasons would be that the player has a favorite weapon, although that is also outside the scope of this analysis. 

We then analyze the `score` variable. It's quantitative and has no missing values. A histogram would be most appropriate here to visualize the score distribution. 

```{r}
ggplot(data = cod_working, mapping = aes(x = score)) +
  geom_histogram(binwidth = 250, fill = 'lightgreen', color = 'black')+
  labs(
    x = 'Score',
    y = 'Number of Games',
    title = 'Call of Duty Score Distribution') +
  theme_bw()
```

